A sensor grid approach for large scale acoustic monitoring

### 2017-07-20 paper/article.tex
ml proof read 3

///

@@

The coder's output bitrate as well as measurement error is computed for both data representations. This metric is composed of band analysis error and the additional loss yielded by quantization before the encoding process. To this aim, third-octave bands are computed with the proposed method, then averaged and compared with the Matlab \textit{ita\_ toolbox}\cite{itatoolbox2017} implementation for the whole signal. We study both "fast" computation with 125~ms windows and no overlap and "slow" 1~s, 50\% overlap analysis to assess the impact of time integration. In both cases the reference is composed of 2~s extracts. The experiment is run for white noise as well as urban environmental recordings. As the analysis frames are of short duration, poor frequency resolution and spectral leakage are important, particularly in lower studied frequencies. For this reason we evaluate the effect of different windowing functions on each band measurement error. The toolbox's implementation shows less than 0.1~dB error when compared with a class 1 sound level meter and as such is considered a sufficient reference. To provide relevant statistics, these quantities are estimated for about 4500 extracts obtained from the UrbanSound8k dataset which we discuss in the next section.

@@

In order to ensure that the proposed scheme allows the recognition of events using state of the art methods, we consider the UrbanSound8k dataset\cite{salamon2014}. It features about 9 hours of urban environmental sounds recordings separated in 8732 wave files ranging from 1 to 4 seconds each. The recordings are labeled in 10 classes (air conditioner, car horn, children playing, dog bark, drilling, enginge idling, gun shot, jackhammer, siren, street music), and distributed in 10 independant folds. A method and baseline results are also provided for the classification task. It is used in all the following experiments with the exception of intelligibility computation for which a small dataset of high-quality voice recordings is preferred. Audio files are resampled 44.1~kHz, normalized and reduced to a single channel.

We evaluate the global loss of information by implementing the four classification models proposed in \cite{salamon2014}: a support vector machine with a radial basis function kernel, a random forest classifier with 500 trees, a decision tree and a k-nearest neighbors classifier with $k = 5$. Values for the SVM parameter $C$ and RBF kernel variance $\sigma^2$ are found with a grid search. We apply a discrete cosine transform to the critical band signal representation to obtain cepstra (known as Mel Frequency Cepstrum Coefficients or MFCC in the first case) of which we conserve the 25 first coefficients, and summarize along time with the mean, variance, skewness, kurtosis, minimum, maximum, median, derivative mean and variance, second order derivative mean and variance. The feature vector is thus comprised of 275 values to reproduce available results and compare with our own. Models are trained for each setup using a 10-fold cross validation method provided by the authors of \cite{salamon2014}, that is, every combination of testing one fold on models trained with the other nine.

@@

The use of an audio recording as an evidence is considered by the forensic phonetics \cite{baldwin1990forensic} domain. It shall be remebered that even recorded with a good quality, such recording remains a weak biometrical indicator \cite{boe2000forensic}. Still, even if the identity of the speaker cannot be asserted, one can attempt to transcribe the spoken utterances. As such, we believe that it is important to demonstrate that the proposed encoding scheme do not allow such thing.

@@

Second, as the speech is mostly concentrated at a given frequency region, a band pass filter from 1 kHz to $~$4-5 kHz could be considered in order to remove the formant information \cite{kent1992acoustic}. However, as some frequency bands are no longer available, the computation of the acoustic indicators is no longer possible.

@@

To evaluate the impact of the frame rate on intelligibility a perceptual test and the evolution of two objective measures are considered. For the evaluation, a dataset of clean speech recordings made of 9 french sentences enunciated by 3 male and 3 female speakers. The recorded sentences are encoded using the proposed approach with analysis frame rates of 50~Hz, 16~Hz, 8~Hz and 4~Hz for a total of 216 audio excerpts. For the perceptual test, each subject is presented with sentences and asked both to transcribe the utterances and rate the global intelligibility between 1 and 5. Due to the low number of sentences, each sentence is tested twice for the same subject but under different encoding setting and speaker identity. 12 subjects of age ranging from 17 to 60 which reported normal hearing participated to the listening test. The test was conducted using a Matlab interface displayed with a desktop computer and audio was played through Beyerdynamics DT 770 headphones. The output level was set by the experimenter at the same level for all subjects.

Concerning the objective measures, we computed the Coherence Speech Intelligibility Index \cite{kates2005} (CSII) and the frequency-weighted segmental SNR \cite{hu2008} (fwSNRseg) by comparing original unaltered samples and recovered audio as those two indicators have been shown to correlate well with perceptual tests \cite{ma2009}, although results are only available for lesser degradations and distortions than the ones we consider in this study.

These metrics require an estimate of the time-domain signal. To do so, a linearly-scaled spectrogram from the band-focused representations is estimated. This estimation can be achieved by multiplying by the scaled transpose of the forward transformation matrix. A loss of resolution increasing with frequency is induced due to the shape of the transformation, as illustrated by Figure~\ref{fig:freq}. Signal phase is approximated by either white noise spectrogram scaling or the Griffin \& Lim algorithm \cite{griffin1984}. The signal is then retrieved using concatenation or overlap-add. When using overlap to compute third-octave bands, one can also avoid framing effects produced by rectangular windowing by convoluting the signal with another windowing function prior to inverse-STFT computation.

@@

The main indicator of performance is the bitrate obtained at the output of the coder. The three varying factors studied are the data word size $q$, the number of bands and the effect of reducing time-resolution by averaging analysis frames. Figure~\ref{fig:bitrate_q} shows estimations of the output bitrate for different values of $q$. To match third-octave bands computation principle where a 125~ms analysis is mandatory, we averaged mel frames over time. Because we use the most common parameters, namely 23.2~ms window with 50\% overlap, the closest achievable rate considering simple averaging is 7.74 frames per second. Third-octave representation on 31 bands yielded an overall higher size than their mel equivalent, here estimated for 30 bands. It however compares with the 40-mel bands representation which is the most used in litterature. This occurence is likely due to the distribution observed by the data prior to Huffman encoding.

@@

A second set of parameters influences directly the time-frequency resolution of the analysis. By choosing a frame rate and number of bands, one can effectively control the size of periodically transmitted data. We evaluate the bitrate for 10 to 40 mel bands and a frame rate from 2 to 10 per second, with fixed $q = 8$. Results are exposed in Figure~\ref{fig:bitrate_mel_avg}. As expected, the bitrate for a given word size $q$ can be modeled as a linear function of the representation dimensions for one second of analysis. Small variations are induced by data distributions on a per-frame level and their effect on the Huffman algorithm.

@@

This experiment aims at assessing potential tradeoffs between bitrate and information. As such, it will be further discussed in the \textit{Event recognition} subsection.\\

The measurement error is computed for third-octave bands and different windowing functions. Results are displayed in Table~\ref{T1}. The error was found to be decreasing with frequency increasing. The study is thus focused on the 20~Hz and 1~kHz bands, respectively the lowest and reference frequencies. The first observation is that using 1~s analysis over 125~ms yields better estimations at 20~Hz. This is expected as it provides a globally better resolution and therefore reduces the impact of spectral leakage effects in this region. When considering white noise the error is low on the whole spectrum, although this effectiveness does not translate to environmental sounds analysis. A factor of this phenomenon is found to be the sound level disparities absent in the first case but omnipresent in the second. For each frame, we computed the difference between the maximum and studied band levels in the reference, and compared it with the observed estimation error. The results show a correlation between the two quantities: when the sound level is low with respect to the peak, the error tends to increase. Furthermore, the closer a significantly higher level is to the studied frequency band, the larger the error is. This is due to framing and as such is particularly important in lower frequencies that suffer the most from it. The window function choice impact is closely related to these concerns. As the bitrate constraint prioritizes no overlap use for a fast analysis, the rectangular window seems a reasonable choice at first with its energy conservation qualities. While it yields the lowest error in high frequency bands, its high spectral leakage effect makes it unprecise in lower frequencies. Non-flat windows better account for this issue but require assuming that the signal is stationnary in a given frame and are thus biased. For slow measurements, the rectangular window impact is less harmful to band analysis due to an increased frequency resolution.\\

\begin{table*}[h]
\centering
\caption{Third-octave bands measurement absolute error in dB for fast and slow computations and four window functions. The proposed method is here compared with evaluations by the \textit{ITA-toolbox} for 2-seconds extracts. }
\label{T1}
\begin{tabular}{|c|c|c|c|c|c|}
\cline{1-6}
\multirow{2}{*}{Analysis} & \multirow{2}{*}{Window} & \multicolumn{2}{|c|}{White noise} & \multicolumn{2}{|c|}{Environmental recordings}\\ \cline{3-6}
 & & 20~Hz & 1~kHz & 20~Hz & 1~kHz\\ \cline{1-6}
\multicolumn{1}{|c}{\multirow{4}{*}{Fast - 125~ms, 0$\%$ overlap}}
 & \multicolumn{1}{|c|}{Rectangular} & 0.94$\pm$0.75 & 0.07$\pm$0.02 & 5.59$\pm$5.64 & 0.21$\pm$0.76\\ \cline{2-6}
\multicolumn{1}{|c}{}
 & \multicolumn{1}{|c|}{Hamming} & 1.02$\pm$0.71\newline & 0.16$\pm$0.14 & 3.34$\pm$3.41 & 0.44$\pm$0.76\\ \cline{2-6}
\multicolumn{1}{|c}{}
 & \multicolumn{1}{|c|}{Gaussian} & 1.04$\pm$0.71 & 0.17$\pm$0.15 & 3.56$\pm$3.63 & 0.48$\pm$0.83\\ \cline{2-6}
\multicolumn{1}{|c}{}
 & \multicolumn{1}{|c|}{Tukey ($\alpha = 0.5$)} & 0.97$\pm$0.75 & 0.11$\pm$0.09 & 4.16$\pm$5.58 & 0.32$\pm$0.80\\ \cline{1-6}
 \multicolumn{1}{|c}{\multirow{4}{*}{Slow - 1~s, 50$\%$ overlap}}
 & \multicolumn{1}{|c|}{Rectangular} & 0.48$\pm$0.36 & 0.08$\pm$0.05 & 1.86$\pm$3.13 & 0.40$\pm$0.55\\ \cline{2-6}
\multicolumn{1}{|c}{}
 & \multicolumn{1}{|c|}{Hamming} & 0.57$\pm$0.56 & 0.09$\pm$0.07 & 2.23$\pm$3.58 & 0.74$\pm$1.61\\ \cline{2-6}
\multicolumn{1}{|c}{}
 & \multicolumn{1}{|c|}{Gaussian} & 0.62$\pm$0.61 & 0.10$\pm$0.08 & 2.48$\pm$4.16 & 0.80$\pm$1.75\\ \cline{2-6}
\multicolumn{1}{|c}{}
 & \multicolumn{1}{|c|}{Tukey ($\alpha = 0.5$)} & 0.53$\pm$0.48 & 0.09$\pm$0.07 & 2.82$\pm$6.64 & 0.61$\pm$1.40\\ \cline{1-6}
\end{tabular}
\end{table*}

>>>>

@@

The coder's output bitrate as well as measurement error is computed for both data representations. This metric is composed of band analysis error and the additional loss yielded by quantization before the encoding process. To this aim, third-octave bands are computed with the proposed method, then averaged and compared to a trusted reference. The chosen reference consiiders the Matlab \textit{ita\_ toolbox} implementation \cite{itatoolbox2017} processed over the whole signal. We study both "fast" computation with 125~ms windows and no overlap and "slow" 1~s, 50\% overlap analysis to assess the impact of time integration. In both cases the sound dataset is the same and is composed of 2~s extracts. The experiment is run for white noise as well as urban environmental recordings. As the analysis frames are of short duration, poor frequency resolution and spectral leakage are expected, particularly in the lower frequencies bands. For this reason we evaluate the effect of the use of different windowing functions on each band measurement error in order to mitigate this loss of precision. The reference implementation shows less than 0.1~dB error when compared with a class 1 sound level meter and as such is considered a sufficient reference. To provide relevant statistics, these quantities are estimated for about 4500 extracts obtained from the UrbanSound8k dataset \cite{salamon2014}.

@@

In order to ensure that the proposed scheme allows the recognition of events using state of the art methods, we consider the results achieved by several regognition systems on the UrbanSound8k dataset\cite{salamon2014}. It features about 9 hours of urban environmental sounds recordings separated in 8732 wave files ranging from 1 to 4 seconds each. The recordings are labeled in 10 classes (air conditioner, car horn, children playing, dog bark, drilling, enginge idling, gun shot, jackhammer, siren, street music), and distributed in 10 independent folds. A method and baseline results are also provided for the classification task. It is used in all the following experiments with the exception of intelligibility computation for which a small dataset of high-quality voice recordings is preferred. Audio files are resampled at 44.1~kHz, normalized and reduced to a single channel.

We evaluate the global loss of information induced by the coder by implementing the four classification models proposed in \cite{salamon2014}: a support vector machine with a radial basis function kernel, a random forest classifier with 500 trees, a decision tree and a k-nearest neighbors classifier with $k = 5$. Values for the SVM parameter $C$ and RBF kernel variance $\sigma^2$ are found with a grid search. We opionaly apply a discrete cosine transform to the critical band signal representation to obtain cepstra (known as Mel Frequency Cepstrum Coefficients or MFCC in the first case) of which we conserve the 25 first coefficients, and summarize along time with the mean, variance, skewness, kurtosis, minimum, maximum, median, derivative mean and variance, second order derivative mean and variance. The feature vector is thus comprised of 275 values to reproduce available results and compare with the data produced by the proposed coding scheme. Models are trained for each setup using the 10-fold cross validation method provided by the authors of \cite{salamon2014}, that is, every combination of testing one fold on models trained with the other nine.

@@

The use of an audio recording as an evidence is considered by the forensic phonetics \cite{baldwin1990forensic} domain. That being said, the readers shall be aware that even recorded with a good quality, such recording remains a weak biometrical indicator \cite{boe2000forensic}. Still, even if the identity of the speaker cannot be asserted, one can attempt to transcribe the spoken utterances. As such, we believe that it is important to demonstrate that the proposed encoding scheme do not allow such thing.

@@

Second, as the speech is mostly concentrated at a given frequency region, a band pass filter from 1 kHz to $~$4-5 kHz could be considered in order to remove the formant information \cite{kent1992acoustic}. However, as some frequency bands are no longer available, the computation of statndard acoustic indicators is no longer possible.

@@

Thus, the impact of the frame rate parameter of the proposed encoding scheme on intelligibility is evaluated. To do so, a perceptual test and the evolution of two objective measures are considered. For the evaluation, a dataset of clean speech recordings made of 9 french sentences enunciated by 3 male and 3 female speakers, for a total of 216 audio excerpts. The recorded sentences are encoded using the proposed approach with analysis frame rates of respectively 50~Hz, 16~Hz, 8~Hz and 4~Hz. For the perceptual test, each subject is presented with sentences and asked both to transcribe the utterances and rate the global intelligibility between 1 and 5. Due to the low number of sentences, each sentence is tested twice for the same subject but under different encoding setting and speaker identity. 12 subjects of age ranging from 17 to 60 which reported normal hearing participated to the listening test. The test was conducted using a Matlab interface displayed with a desktop computer and audio was played through \textit{Beyerdynamics DT 770} headphones. The output level was set by the experimenter at the same level for all subjects.

Concerning the objective measures, the Coherence Speech Intelligibility Index \cite{kates2005} (CSII) and the frequency-weighted segmental SNR \cite{hu2008} (fwSNRseg) are considered. They both compare original unaltered samples and recovered audio as those two indicators have been shown to correlate well with perceptual tests \cite{ma2009}, although results are only available for lesser degradations and distortions than the ones we consider in this study.

These metrics require an estimate of the time-domain signal. To do so, a linearly-scaled spectrogram from the band-focused representations is estimated. This estimation can be achieved by multiplying by the scaled transpose of the forward transformation matrix. A loss of resolution increasing with frequency is induced due to the shape of the transformation, as illustrated by Figure~\ref{fig:freq}. Signal phase is then approximated by either white noise spectrogram scaling or the Griffin \& Lim algorithm \cite{griffin1984}. The signal is then retrieved using concatenation or overlap-add depending on the overlap setting. When using overlap to compute third-octave bands, one can also avoid framing effects produced by rectangular windowing by convoluting the signal with another windowing function prior to inverse-STFT computation.

@@

The main indicator of performance is the bitrate obtained at the output of the coder. The three varying factors studied are the data word size $q$, the number of bands and the effect of reducing time-resolution by averaging analysis frames. Figure~\ref{fig:bitrate_q} shows estimations of the output bitrate for different values of $q$. To match third-octave bands computation principle where a 125~ms analysis is mandatory, we averaged mel frames over time. Because we use the most common parameters, namely 23.2~ms window with 50\% overlap, the closest achievable rate considering simple averaging is 7.74 frames per second. Third-octave representation on 31 bands yielded an overall higher size than their mel equivalent, here estimated for 30 bands. It however compares with the 40-mel bands representation which is the most used features in litterature. This hugher bitrate is likely due to the distribution observed by the data prior to Huffman encoding.

@@

A second set of parameters influences directly the time-frequency resolution of the analysis. By choosing a frame rate and number of bands, one can effectively control the size of periodically transmitted data. We evaluate the bitrate for 10 to 40 mel bands and a frame rate from 2 to 10 per second, with fixed $q = 8$. Results are shown in Figure~\ref{fig:bitrate_mel_avg}. As expected, the bitrate for a given word size $q$ can be modeled as a linear function of the representation dimensions for one second of analysis. Small variations are induced by data distributions on a per-frame level and their impact on the Huffman algorithm.

@@

This experiment aims at an initial assessement of potential tradeoffs between bitrate and preservation of valuable information. As such, it will be further discussed in the \textit{Event recognition} subsection.\\

The measurement error is computed for third-octave bands and different windowing functions. Results are displayed in Table~\ref{T1}. The error was found to be decreasing with respect to an increase of frequency. The study is thus focused on the 20~Hz and 1~kHz bands, respectively the lowest and reference frequencies. The first observation is that using 1~s analysis over 125~ms yields better estimations at 20~Hz. This is expected as it provides a globally better resolution and therefore reduces the impact of spectral leakage effects in this region.

Closer inspection show that when considering white noise the error is low on the whole spectrum. This effectiveness does not translate though to environmental sounds analysis. One inducing factor for this phenomenon is the sound level disparities absent in the first case but omnipresent in the second. Let us consider, for each frame, the difference between the maximum and the remaining band levels in the reference, and compared it with the observed estimation error. The results show a correlation between the two quantities: when the average band level is low with respect to the peak, the error tends to increase. Furthermore, the closer a significantly higher level is to the considered frequency band, the larger the error is. This is due to framing that causes spectral leakage. This is found to be particularly important in lower frequencies.

To mitigate this issue, the choice of the window function shall be studied further. As the bitrate constraint prioritizes no overlap use for a fast analysis, the rectangular window seems a reasonable choice at first with its energy conservation qualities. While it yields the lowest error in high frequency bands, its high spectral leakage effect makes it unprecise for lower frequency regions. Non-flat windows better account for this issue but require assuming that the signal is stationnary in a given frame and are thus biased. For slow measurements, the rectangular window impact is less harmful to band analysis due to an increased frequency resolution.\\

\begin{table*}[h]
\centering
\caption{Third-octave bands measurement absolute error in dB for fast and slow computations and four window functions. The proposed method is here compared with evaluations by the \textit{ITA-toolbox} for 2-seconds exerpts. }
\label{T1}
\begin{tabular}{|c|c|c|c|c|c|}
\cline{1-6}
\multirow{2}{*}{Analysis} & \multirow{2}{*}{Window} & \multicolumn{2}{|c|}{White noise} & \multicolumn{2}{|c|}{Environmental recordings}\\ \cline{3-6}
 & & 20~Hz & 1~kHz & 20~Hz & 1~kHz\\ \cline{1-6}
\multicolumn{1}{|c}{\multirow{4}{*}{Fast - 125~ms, 0$\%$ overlap}}
 & \multicolumn{1}{|c|}{Rectangular} & 0.94$\pm$0.75 & 0.07$\pm$0.02 & 5.59$\pm$5.64 & 0.21$\pm$0.76\\ \cline{2-6}
\multicolumn{1}{|c}{}
 & \multicolumn{1}{|c|}{Hamming} & 1.02$\pm$0.71\newline & 0.16$\pm$0.14 & 3.34$\pm$3.41 & 0.44$\pm$0.76\\ \cline{2-6}
\multicolumn{1}{|c}{}
 & \multicolumn{1}{|c|}{Gaussian} & 1.04$\pm$0.71 & 0.17$\pm$0.15 & 3.56$\pm$3.63 & 0.48$\pm$0.83\\ \cline{2-6}
\multicolumn{1}{|c}{}
 & \multicolumn{1}{|c|}{Tukey ($\alpha = 0.5$)} & 0.97$\pm$0.75 & 0.11$\pm$0.09 & 4.16$\pm$5.58 & 0.32$\pm$0.80\\ \cline{1-6}
 \multicolumn{1}{|c}{\multirow{4}{*}{Slow - 1~s, 50$\%$ overlap}}
 & \multicolumn{1}{|c|}{Rectangular} & 0.48$\pm$0.36 & 0.08$\pm$0.05 & 1.86$\pm$3.13 & 0.40$\pm$0.55\\ \cline{2-6}
\multicolumn{1}{|c}{}
 & \multicolumn{1}{|c|}{Hamming} & 0.57$\pm$0.56 & 0.09$\pm$0.07 & 2.23$\pm$3.58 & 0.74$\pm$1.61\\ \cline{2-6}
\multicolumn{1}{|c}{}
 & \multicolumn{1}{|c|}{Gaussian} & 0.62$\pm$0.61 & 0.10$\pm$0.08 & 2.48$\pm$4.16 & 0.80$\pm$1.75\\ \cline{2-6}
\multicolumn{1}{|c}{}
 & \multicolumn{1}{|c|}{Tukey ($\alpha = 0.5$)} & 0.53$\pm$0.48 & 0.09$\pm$0.07 & 2.82$\pm$6.64 & 0.61$\pm$1.40\\ \cline{1-6}
\end{tabular}
\end{table*}
